On the minimum leaf conductance: its role in models of plant water use, and ecological and environmental controls

### 2018-07-18 manuscript.Rmd
final minor edits for copyeditor

\
Remko A. Duursma^1^, Chris J. Blackman^1^, Rosana Lopéz^1,3^, Nicolas K. Martin-StPaul^2^, Hervé Cochard^3^, Belinda E. Medlyn^1^

@@

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories: conductance of the cuticle only, conductance of detached leaves, and gas-exchange measurements during conditions leading to presumed stomatal closure. The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss some important aspects of cuticular conductance in Section III.

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain. In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S2). It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].

@@

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S1 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S1 for full details of the compilation). The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low\ PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low\ A~ leaf conductance during periods of very low photosynthesis. See text for data sources and methods.

@@

\
There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008]. The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S2), with a typical decrease in the order of 30-40%. In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978]. We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig. 3). All 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details). This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.

@@

\
Finally we mention the striking effect of leaf age on g~min~. We have summarized five studies in Fig. S3, all of which reported an increased g~min~ for older leaves for the majority of species studied. The data reported by @jordan2007 are particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca. 10 years. If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section III). Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007. In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.

@@

\
**Figure 4**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in an elevated temperature treatment (+3 °C) [see @drake2018]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model).

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero. The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of @ball1988, as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is usually estimated from a regression of Eq. 1 with leaf gas exchange data, but as we will show, this approach may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

@@

\
**Figure 7**. Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. The right panel visualizes the correlation between estimated slope (g~1~) and intercept (g~0~) of the regression. The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

@@

**Methods S1** Literature compilation of g~min~ estimates

**Methods S2** Methodological issues with measurement of g~min~

>>>>

\
Remko A. Duursma^1^ (ORCID: 0000-0002-8499-5580), Christopher J. Blackman^1^, Rosana Lopéz^1,3^, Nicolas K. Martin-StPaul^2^ (ORCID: 0000-0001-7574-0108), Hervé Cochard^3^ (ORCID: 0000-0002-2727-7072), Belinda E. Medlyn^1^ (ORCID: 0000-0001-5728-9827)

@@

Keywords: cuticular conductance, drought tolerance, ecosystem modelling, plant water relations, stomatal conductance model

@@

\
Measurements of minimum conductance, after stomatal closure is either induced or assumed, can be broadly divided into the following categories: conductance of the cuticle only, conductance of detached leaves, and gas-exchange measurements during conditions leading to presumed stomatal closure (Table 1). The absolute minimum conductance attained by leaves is that through the cuticle only. Measurements of cuticular conductance (g~cuti~) are typically made on isolated non-stomatal cuticles [@riederer2001], via special gas exchange techniques [@boyer1997], or mass loss of detached leaves by sealing the side with stomata [@kerstiens1996]. We further discuss some important aspects of cuticular conductance in Section III.

\
The minimum conductance of intact leaves is typically measured by mass loss of detached leaves (MLD). In this method, leaves are detached and leaf mass is monitored over time as the leaf dries out. Early work by @hygen1951 showed that after a leaf is detached, initial water loss rates are high, but after some time a constant low rate is achieved. From this minimum transpiration rate (sometimes described as the 'residual transpiration', or 'epidermal transpiration'), the minimum conductance (g~min~) can be estimated using measured vapour pressure deficit [@sinclair1986]. Although this method resembles conditions that plants may experience during a dry-down or periods of extreme stress, some uncertainties regarding the methods for measurement remain. In our own work, we have found that in some leaves, water loss rate increases (rather than decreases) some time after leaf detachment (see Methods S1). It is likely that the relatively rapid dry-down sometimes causes artefacts, and must be carefully avoided [see also @heinsoo1998].

@@

\
Measurements of conductance of isolated, non-stomatal cuticles (g~cuti~) were taken from the compilations by @kerstiens1996 and @schuster2017 (both also include g~min~ data but these were not used since they included no metadata). Data on minimum conductance (g~min~) from MLD was compiled from 40 original papers (see Methods S2 for description and references). For the compilation, we returned to all original papers mentioned in the review by @kerstiens1996, and added many newer sources (see Methods S2 for full details of the compilation). The database includes a total of 221 species (136 Angiosperm, 49 Gymnosperm, 1 Pteridophyte), from 57 taxonomic families, with woody species making up roughly 2/3 of the data.

@@

\
**Figure 1**. Comparison of various estimates of the (presumed) minimum conductance. Error bars are 95\% confidence intervals. Grey dots are the original data (but a few data points occur outside the figure range). Different letters denote significant differences (at $\alpha$ = 0.05). g~cuti~ : conductance of isolated cuticles; g~min~ : minimum conductance measured with mass loss of detached leaves; g~dark~ : leaf conductance during the night or after dark adaptation; g~low\ PAR~ leaf conductance during low light (photosynthetically active radiation, PAR); g~low\ A~ leaf conductance during periods of very low photosynthesis. See Section II for data sources and methods.

@@

\
There is a general tendency for a decreased g~min~ in plants acclimated to drought stress [@james2008]. The magnitude of the decrease of g~min~ with acclimation to drought stress varied from -4 to -70% (across 10 studies, see Table S1), with a typical decrease in the order of 30-40%. In each of the studies summarized, plants were grown in well-watered or droughted conditions, and in one case a difference in g~min~ was demonstrated after just four days of drought exposure [@bengtson1978]. We also demonstrate drought acclimation via a case study on 11 *Hakea* species grown in two watering treatments (Fig. 3). All 11 species showed a decrease in g~min~ in the drought treatment (see Methods S3 for experimental details). This significant change in g~min~ with drought acclimation is likely an important component of the overall drought hardening of plants.

@@

\
Finally we mention the striking effect of leaf age on g~min~. We have summarized five studies in Fig. S1, all of which reported an increased g~min~ for older leaves for the majority of species studied. The data reported by @jordan2007 are particularly impressive, as g~min~ in the woody shrub *Agastachys odorata* gradually increased with leaf age up to ca. 10 years. If this effect is due to properties of the cuticle alone, a possible explanation for the increase in g~min~ is the continued exposure to wind, rain, and abrasives, which have been shown to damage the cuticle and increase its conductance (see discussion in Section III). Another possibility is that the contribution of stomata to g~min~ increases with leaf age, as reported by @jordan2007. In plants that maintain several cohorts of leaves, finding an appropriate value of g~min~ for use in models will have to take into account the leaf age effect.

@@

\
**Figure 4**. Minimum conductance (g~min~) measured with mass loss of detached leaves on *Eucalyptus parramattensis* grown in whole-tree chambers. Trees were grown following ambient conditions, or in an elevated temperature treatment (+3 °C) [see @drake2018]. Measurements of g~min~ were carried out at five different temperatures on replicate leaves. There was no consistent effect of measurement temperature on g~min~, but leaves of trees grown in elevated temperature showed on average a 45% reduction in g~min~ (p < 0.01, linear mixed-effects model). Error bars denote 95% confidence intervals.

@@

\
Where g~1~ is a 'slope' parameter, A~n~ the leaf net photosynthetic rate, C~a~ atmospheric CO~2~ concentration, f(D) is some function of the vapour pressure deficit D (or relative humidity in case of the model of @ballberry1987, see @damour2010 for a list of functions), and g~0~ is the value of g~s~ when A~n~ is zero. The g~1~ parameter is directly related to the water-use efficiency: large values of g~1~ indicate low water-use efficiency. In this paper, we refer to Eq. 1 as 'the Ball-Berry model', thus including all model formulations that include different f(D) functions besides the one proposed by Ball & Berry (1987). Much attention has been paid to quantifying and interpreting variation in the g~1~ parameter [@lin2015; @miner2017; @prentice2014; @medlyn2011], but the g~0~ parameter has been studied in much less detail. For example, a recent comprehensive review of stomatal conductance models did not mention g~0~ or any similar minimum conductance [@damour2010]. The original description of Eq. 1 as published by @ballberry1987 did not include an intercept term. It was first introduced in the unpublished manuscript of Ball & Berry (1988, available on request), as an 'intercept' without further discussion of the role it plays in the model. In practice, g~0~ is usually estimated from a regression of Eq. 1 with leaf gas exchange data, but as we will show, this approach may lead to inaccurate values, as pointed out by @barnard2013 but not explained in detail.

@@

\
**Figure 7**. (a) Example linear regression of stomatal conductance (g~s~) against a combination of terms, in this case the linearized version of the model of Medlyn et al. (2011) applied to a leaf gas exchange dataset of @martinstpaul2012. This example shows a very good fit between g~s~ and the stomatal index, and was selected from the database of Lin et al. (2015). The solid line is regression line, and the shaded area a 95% confidence interval for the mean. (b) The correlation between estimated slope (g~1~) and intercept (g~0~) of the regression shown in panel (a). The dotted ellipse is a bivariate 95% confidence interval for slope and intercept. The symbols represent 1000 bootstrap samples of the coefficients.

@@

**Methods S1** Methodological issues with measurement of g~min~

**Methods S2** Literature compilation of g~min~ estimates
